---
title: "Perturbation-based benchmark"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{perturbBench}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Peturbation-based benchmark
## Introduction
This benchmark is based on experiments where specific kinases are expected to 
have an increased or decreased activity after being perturbed in 
various cell line experiments. 

We provide the log fold changes of phosphorylation sites after the perturbation
which can be used to infer changes in kinase activities using different methods
or kinase-substrate libraries. 

The methods are then evaluated based on whether they are able to recapitulate
the peturbed kinases based on their inferred activities. 

## Getting set up
We first load the required packages to run the benchmark.
```{r setup}
library(benchmarKIN)
library(dplyr)
```

Since we call the get_performances function from the decoupler-py package
we also need to set up reticulate.

```{r reticulate}
library(reticulate)

# Use reticulate to install the Python package
py_install("git+https://github.com/saezlab/decoupler-py.git", pip = TRUE)

# If you are using a conda environment make sure to use the correct path for your
# python version and install the decoupler-py package directly into your environment

# use_python("path-to-python", required = TRUE)  #you can find the correct path by activating your environment and typing "which python" into the terminal
# py_install("git+https://github.com/saezlab/decoupler-py.git", pip = TRUE, envname = "path-to-environment")
```

## Perturbation data
The perturbation data can be extracted using the load_perturbData() function.
This data frame contains the logFC for the perturbation collection previously
published by Hernandez-Armenta et al. and the perturbation experiments from
Hijazi et al.

```{r data}
mat <- load_perturbData()
mat[1:5, 20:23]
```
Additional information about each experiment can be found in the meta data. This
contains information about the targeted kinase(s), the PMID, etc.

```{r metadata}
meta <- load_meta()
head(meta)
```
## Kinase activity inference
This data can be used to test any method for kinase activity inference. 

We will use the z-score (as implemented by RoKAI) and the PhosphoSitePlus kinase-
substrate library as in example.

### Kinase-substrate library
For that we have already processed a version of PhosphoSitePlus (accessed: 19/04/2023)
that can be mapped to our phosphorylation site ids.

```{r phosphositeplus}
head(phosphositeplus)
```
After mapping the phosphorylation sites we can bring it into the right format to
run the run_zscore function.

```{r map_ppsp}
phosphositeplus$target <- paste(phosphositeplus$target, phosphositeplus$target_protein, phosphositeplus$position, sep = "|")

ppsp <- phosphositeplus %>%
  dplyr::select(source, target, mor) %>%
  dplyr::distinct()

head(ppsp)
```

### Activity inference using the z-score 
The z-score calculates a score for each kinase by aggregating the change in
abundance of the direct targets in relation to changes in the non-targets.

```{r activity}
act_scores <- run_zscore(mat = mat, network = ppsp)

act_scores[1:5, 1:5]
```
## Perturbation benchmark
### Area under the receiver operator curve
The inferred kinase activities can now be evaluated using the run_perturbBench()
function. Hereby, it is important to note that the kinase symbol needs to match
with the meta data (here gene symbols).

```{r performance}
performance <- run_perturbBench(act = act_scores, meta = meta, method_id = "zscore_ppsp")
```

We can then plot the areas under the receiver operator curve (AUROCs) and can 
compare this with other methods or prior knowledge resources.
```{r plot}
auroc_p <- performance %>%
  ggplot2::ggplot(ggplot2::aes(x=method, y=auroc)) +
  ggplot2::geom_boxplot(outlier.size=1, lwd=0.6) +
  ggplot2::geom_hline(yintercept = 0.5)

auroc_p
```

### Area under the receiver operator curve
Additionally we can calculate the rank and scaled rank of the perturbed kinase(s)
in their respective experiments. A lower rank/scaled rank indicates a better
performance of the method

```{r scaled_rank}
rank <- run_rank(act = act_scores, meta = meta)
median(rank$scaled_rank)
```

